---
title: "thesisWorkspace"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(tidyverse)
library(knitr)
library(readr)
library(party)
library(modelr)
library(xgboost)
```


```{r}
df <- read_csv("./data/combined-pre.csv") %>% select(Div, Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HTHG, HTAG, HTR, HS, AS, HST, AST, HF, AF, HY, AY, HR, AR, B365H, B365D, B365A)

# need to bring in 2019-2020, 2020-2021, and 2021-2022 data in sperately bc they did not combine correctly

df1 <- read_csv("./data/2019-2020.csv") %>% select(Div, Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HTHG, HTAG, HTR, HS, AS, HST, AST, HF, AF, HY, AY, HR, AR, B365H, B365D, B365A)

df2 <- read_csv("./data/2020-2021.csv") %>% select(Div, Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HTHG, HTAG, HTR, HS, AS, HST, AST, HF, AF, HY, AY, HR, AR, B365H, B365D, B365A)

df3 <- read_csv("./data/2021-2022.csv") %>% select(Div, Date, HomeTeam, AwayTeam, FTHG, FTAG, FTR, HTHG, HTAG, HTR, HS, AS, HST, AST, HF, AF, HY, AY, HR, AR, B365H, B365D, B365A)

# La Liga match ups from the 2008-2009 season up until the 2021-2022 season
df <- rbind(df, df1, df2, df3)
```

## La Liga data
```{r}
# 08/09 season to 21/22 
classico <- df %>% filter(HomeTeam=="Real Madrid" & AwayTeam=="Barcelona" | AwayTeam=="Real Madrid" & HomeTeam=="Barcelona")

w <- df %>% filter(HomeTeam=="Celta") %>% select(-Date, -Div, -FTHG, -FTAG)
w

w_f <- df %>% filter(HomeTeam=="Celta") %>% select(-Date, -Div, -FTHG, -FTAG, -AwayTeam, -HomeTeam)
w_f

write_csv(classico, "./data/classico.csv")

classico_new <- read_csv("./data/Copy_of_classico.csv")
classico_for_xgb <- classico_new %>% select(-Date, -Div, -FTHG, -FTAG, -HomeTeam, -AwayTeam)
classico_no_chars <- classico_new %>% select(-Date, -Div, -FTHG, -FTAG)
```

```{r}
xrf <- classico_new %>% mutate(HalfTimeGoals = HTHG - HTAG,
                                            ShotCount = HS - AS,
                                            ShotsTargetCount = HST - AST,
                                            FoulsCommited = HF - AF,
                                            YellowCount = HY - AY,
                                            RedCount = HR - AR,
                                            Rank = `H-Rank` - `A-Rank`,
                                            CoachTenure = `HC. Tenure` - `AC. Tenure`,
                                            CoachAge = `HC. Age` - `AC. Age`
                                            ) %>% select(-Div, -Date, -FTHG, -FTAG, -HTHG, -HTAG, -HS, -AS, -HST, -AST, -HF, -AF, -HY, -AY, -HR, -AR, -`H-Rank`, -`A-Rank`, -`HC. Tenure`, -`AC. Tenure`, -`HC. Age`, -`AC. Age`, -HomeTeam, -AwayTeam)
xrf

yxgb <- xrf
```

```{r}
w_f[,] <- lapply(w_f[,], as.factor)
#dim(classico_new)
#dim(w)
classico_no_chars[,] <- lapply(classico_no_chars[,], as.factor)
xrf[,] <- lapply(xrf[,], as.factor)
```




```{r}
# NORMAL DATA

set.seed(2) #seed for reproducibility
index <- round(nrow(classico_no_chars) * 0.25, digits=0)
test_indices <- sample(1:nrow(classico_no_chars), index)
train_set <- classico_no_chars[-test_indices,] #80% training set
test_set <- classico_no_chars[test_indices,]  #20% test set

set.seed(2)
index_xgb <- round(nrow(classico_for_xgb) * 0.25, digits=0)
test_indices_xgb <- sample(1:nrow(classico_for_xgb), index_xgb)
train_set_xgb <- classico_for_xgb[-test_indices_xgb,] #80% training set
test_set_xgb <- classico_for_xgb[test_indices_xgb,]  #20% test set

# CELTA 

set.seed(24) #seed for reproducibility
indexW <- round(nrow(w_f) * 0.25, digits=0)
test_indicesW <- sample(1:nrow(w_f), indexW)
train_setW <- w_f[-test_indicesW,] #80% training set
test_setW <- w_f[test_indicesW,]  #20% test set

set.seed(24)
index_xgbW <- round(nrow(w) * 0.25, digits=0)
test_indices_xgbW <- sample(1:nrow(w), index_xgbW)
train_set_xgbW <- w[-test_indices_xgbW,] #80% training set
test_set_xgbW <- w[test_indices_xgbW,]  #20% test set

# CHANGED VARIABLES (NORMAL DATA)

set.seed(12) #seed for reproducibility
xrf_index <- round(nrow(xrf) * 0.25, digits=0)
xrf_test_indices <- sample(1:nrow(xrf), xrf_index)
xrf_train_set <- xrf[-xrf_test_indices,] #80% training set
xrf_test_set <- xrf[xrf_test_indices,]  #20% test set

set.seed(12) #seed for reproducibility
yxgb_index <- round(nrow(yxgb) * 0.25, digits=0)
yxgb_test_indices <- sample(1:nrow(yxgb), yxgb_index)
yxgb_train_set <- yxgb[-yxgb_test_indices,] #80% training set
yxgb_test_set <- yxgb[yxgb_test_indices,]  #20% test set
```

```{r}
set.seed(123)
cforest_model <- cforest(FTR~., 
                         data = xrf_train_set,
                         control = cforest_control(ntree=200, 
                                                   mtry=5
                                                   #minsplit = 6,
                                                   #minbucket = 2
                                                   )
                         )
cforest_model
set.seed(1)
x <- varimp(cforest_model, conditional = F) %>% sort(decreasing = T)
data.frame(x)
```

```{r}
#train_set$FTR
#test_set$FTR
```

```{r}
#predict(cforest_model, newdata=train_set)
#predict(cforest_model, newdata=test_set)
#predict(cforest_model, newdata=test_set, type="prob")
```

```{r}
#library(vip)
#set.seed(1)
#vip(cforest_model, num_features = 30)
```

```{r}
set.seed(1)
rf_pred_test <- predict(cforest_model, newdata=xrf_test_set)
rf_pred_test_probs <- predict(cforest_model, newdata=xrf_test_set, type="prob")
rf_pred_test_list <- as.data.frame(rf_pred_test)
rf_pred_test_list
```

```{r}
# accuracy measures for cforest
rf_test_accuracy <- xrf_test_set %>% mutate(rf_pred_test = rf_pred_test)
rf_test_accuracy<- rf_test_accuracy %>% mutate(correct = ifelse(rf_pred_test == FTR,1,0))
rf_accuracy <- sum(rf_test_accuracy$correct) / length(rf_test_accuracy$correct)
#rf_accuracy
```


```{r}
#test_set
```


```{r}
yxgb_test_set["HTR"][yxgb_test_set["HTR"] == "H"] <- "1"
yxgb_test_set["HTR"][yxgb_test_set["HTR"] == "D"] <- "2"
yxgb_test_set["HTR"][yxgb_test_set["HTR"] == "A"] <- "3"

yxgb_test_set["FTR"][yxgb_test_set["FTR"] == "H"] <- "1"
yxgb_test_set["FTR"][yxgb_test_set["FTR"] == "D"] <- "2"
yxgb_test_set["FTR"][yxgb_test_set["FTR"] == "A"] <- "3"

test_set_xgb_noName <- yxgb_test_set #%>% select(-HomeTeam, -AwayTeam)

test_set_xgb_noName[,] <- lapply(test_set_xgb_noName[,], as.numeric)

test_set_xgb_noName_noFTR <- test_set_xgb_noName %>% select(-FTR)


yxgb_train_set["HTR"][yxgb_train_set["HTR"] == "H"] <- "1"
yxgb_train_set["HTR"][yxgb_train_set["HTR"] == "D"] <- "2"
yxgb_train_set["HTR"][yxgb_train_set["HTR"] == "A"] <- "3"

yxgb_train_set["FTR"][yxgb_train_set["FTR"] == "H"] <- "1"
yxgb_train_set["FTR"][yxgb_train_set["FTR"] == "D"] <- "2"
yxgb_train_set["FTR"][yxgb_train_set["FTR"] == "A"] <- "3"

train_set_xgb_noName <- yxgb_train_set #%>% select(-HomeTeam, -AwayTeam)

train_set_xgb_noName[,] <- lapply(train_set_xgb_noName[,], as.numeric)

train_set_xgb_noName_noFTR <- train_set_xgb_noName %>% select(-FTR)

train_set_xgb_noName_noFTR

# as.matrix(train_set_xgb_noName_noFTR)
```

Gradient boosting iterates over errors from several "weak" classifying algorithms. These weak classifiers are
called so because they are not exceptionally accurate alone. However, gradient boosting builds a model by evaluating the errors each weak classifier makes. By iterating over these errors, the weak learners contribute to the creation of a stronger model, which culminates in our XGBoost model.

```{r}
set.seed(123)
xg_train <- xgboost(as.matrix(train_set_xgb_noName_noFTR), 
                    train_set_xgb_noName$FTR, 
                    nthread = 2, 
                    nrounds = 15,                               
                    subsample = 0.5,
                    objective = "multi:softmax",
                    num_class = 4)
```

```{r}
importance <- xgb.importance(model = xg_train) 
importance
importance %>% ggplot() + geom_bar(mapping  = aes(x= importance$Feature, y = importance$Gain), stat = "identity") + coord_flip()
```


```{r}
pred <- predict(xg_train, as.matrix(train_set_xgb_noName_noFTR))
pred_test <- predict(xg_train, as.matrix(test_set_xgb_noName_noFTR))
train_prediction_list <- as.data.frame(pred)
test_prediction_list <- as.data.frame(pred_test)

train_prediction_list
test_prediction_list

test_set_xgb_noName$FTR
```

```{r}
xgb_test_accuracy <- test_set_xgb_noName %>% mutate(pred_test = pred_test)
xgb_test_accuracy<- xgb_test_accuracy %>% mutate(correct = ifelse(pred_test == FTR,1,0))
xgb_accuracy <- sum(xgb_test_accuracy$correct) / length(xgb_test_accuracy$correct)
```


```{r}
rf_accuracy
xgb_accuracy
```

Avg Accuracy For 25 Different train/test sets using CHANGED variables!
(el classico)
minsplit = default
minbucket = default
Dim: [1] 45 19
```{r}
seed_num5 <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25)

list_rf_accuracy5 <- c(0.6363636, 0.2727273, 0.7272727, 0.4545455, 0.8181818, 0.6363636, 0.4545455, 0.4545455, 0.7272727, 0.6363636, 0.6363636, )

list_xgb_accuracy5 <- c(0.7272727, 0.5454545, 0.6363636, 0.6363636, 0.5454545, 0.6363636, 0.4545455, 0.5454545, 0.5454545, 0.4545455, 0.6363636, )

df_acc5 <- data.frame(seed_num5, list_rf_accuracy5, list_xgb_accuracy5) 
rf_mean_acc5 <- mean(df_acc5$list_rf_accuracy5)
xgb_mean_acc5 <- mean(df_acc5$list_xgb_accuracy5)

rf_mean_acc5
xgb_mean_acc5
```


Avg Accuracy For 25 Different train/test sets using ALL variables
(el classico)
minsplit = 6
minbucket = 2
Dim [1] 45 34
```{r}
seed_num1 <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25)

list_rf_accuracy1 <- c(0.6363636, 0.4545455, 0.6363636, 0.3636364, 0.5454545, 0.6363636,  0.3636364, 0.4545455, 0.5454545, 0.7272727, 0.7272727, 0.7272727, 0.5454545, 0.3636364, 0.6363636, 0.6363636, 0.5454545, 0.5454545, 0.5454545, 0.5454545, 0.6363636, 0.4545455, 0.7272727, 0.5454545, 0.3636364)

list_xgb_accuracy1 <- c(0.9090909, 0.4545455, 0.7272727, 0.7272727, 0.5454545, 0.6363636, 0.6363636, 0.5454545, 0.6363636, 0.6363636, 0.6363636, 0.6363636, 0.5454545, 0.8181818, 0.7272727, 0.7272727, 0.7272727, 0.7272727, 0.8181818, 0.7272727, 0.7272727, 0.4545455, 0.6363636, 0.6363636, 0.3636364)

df_acc1 <- data.frame(seed_num1, list_rf_accuracy1, list_xgb_accuracy1) 
rf_mean_acc1 <- mean(df_acc1$list_rf_accuracy1)
xgb_mean_acc1 <- mean(df_acc1$list_xgb_accuracy1)

rf_mean_acc1
xgb_mean_acc1
```

Avg Accuracy For 25 Different train/test sets using ALL variables
(el classico)
Default minsplit
Default minbucket
DIm: [1] 45 34
```{r}
seed_num4 <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25)

list_rf_accuracy4 <- c(0.5454545, 0.5454545, 0.6363636, 0.3636364, 0.5454545, 0.4545455, 0.4545455, 0.4545455, 0.2727273, 0.7272727, 0.7272727, 0.6363636, 0.4545455, 0.3636364, 0.3636364, 0.5454545, 0.3636364, 0.3636364, 0.4545455, 0.6363636, 0.5454545, 0.3636364, 0.5454545, 0.4545455, 0.5454545)

list_xgb_accuracy4 <- c(0.9090909, 0.4545455, 0.7272727, 0.7272727, 0.5454545, 0.6363636, 0.6363636, 0.5454545, 0.6363636, 0.6363636, 0.6363636, 0.6363636, 0.5454545, 0.8181818, 0.7272727, 0.7272727, 0.7272727, 0.7272727, 0.8181818, 0.7272727, 0.7272727, 0.4545455, 0.6363636, 0.6363636, 0.3636364)

df_acc4 <- data.frame(seed_num4, list_rf_accuracy4, list_xgb_accuracy4) 
rf_mean_acc4 <- mean(df_acc4$list_rf_accuracy4)
xgb_mean_acc4 <- mean(df_acc4$list_xgb_accuracy4)

rf_mean_acc4
xgb_mean_acc4
```

Avg Accuracy For 25 Different train/test sets using base variables
(Hometeam == "Celta")
minsplit = 6
minbucket = 2
Dim: [1] 184  19
```{r}
seed_num3 <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25)

list_rf_accuracy3 <- c(0.6956522, 0.6304348, .5869565, 0.5, 0.6956522, 0.5434783, 0.6521739, 0.5217391, 0.6521739, 0.5869565, 0.5869565, 0.6304348, 0.6304348, 0.6086957, 0.5217391, 0.4782609, 0.5434783, 0.6086957, 0.6521739, 0.6086957, 0.4782609, 0.5217391, 0.5869565, 0.5434783, 0.5434783)

list_xgb_accuracy3 <- c(0.6304348, 0.6086957, 0.5217391, 0.5434783, 0.6086957, 0.5652174, 0.6304348, 0.5217391, 0.6956522, 0.4782609, 0.5, 0.5217391, 0.673913, 0.6086957, 0.4782609, 0.5, 0.5652174, 0.5434783, 0.5869565, 0.6521739, 0.5, 0.5619565, 0.5217391, 0.4130435, 0.4565217)

df_acc3 <- data.frame(seed_num3, list_rf_accuracy3, list_xgb_accuracy3) 
rf_mean_acc3 <- mean(df_acc3$list_rf_accuracy3)
xgb_mean_acc3 <- mean(df_acc3$list_xgb_accuracy3)

rf_mean_acc3
xgb_mean_acc3
```


Avg Accuracy For 25 Different train/test sets using base variables
(Hometeam == "Celta")
Default minsplit
Default minbucket
Dim [1] 184  19
```{r}
seed_num2 <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25)

list_rf_accuracy2 <- c(0.7173913, 0.6086957, 0.5652174, 0.4782609, 0.7391304, 0.5, 0.5652174, 0.7173913, 0.6304348, 0.6086957, 0.5652174, 0.6304348, 0.6304348, 0.5869565, 0.5217391, 0.4130435, 0.5869565, 0.673913, 0.6086957, 0.6304348, 0.4782609, 0.5, 0.5652174, 0.5652174, 0.4782609)

list_xgb_accuracy2 <- c(0.6304348, 0.6086957, 0.5217391, 0.5434783, 0.6086957, 0.5652174, 0.3913043, 0.6304348, 0.6956522, 0.4782609, 0.5, 0.5217391, 0.673913, 0.6086957, 0.4782609, 0.5, 0.5652174, 0.5434783, 0.5869565, 0.6521739, 0.5, 0.5652174, 0.5217391, 0.4130435, 0.4565217)

df_acc2 <- data.frame(seed_num2, list_rf_accuracy2, list_xgb_accuracy2) 
rf_mean_acc2 <- mean(df_acc2$list_rf_accuracy2)
xgb_mean_acc2 <- mean(df_acc2$list_xgb_accuracy2)

rf_mean_acc2
xgb_mean_acc2
```

